Surface programmable materials

### 2024-08-22 paper/main.tex
modified structure of the paper

\subsection{The direction of computation}\label{sec:direction}

@@

The computation contains the following steps: (1). Initialize the in-surface configuration. By removing
some atoms on the in-surface. (2). Connect the in-surface/out-surface to
external heat sources at temperature $T_1 < T_2$, respectively. We
also require that the energy gap between the ground state and the first
excited state of the Hamiltonian to be $\Delta E< T_1$. (3). Lower the
temperature of the heat sources ``slowly'' to cool the system to the
ground state of the Hamiltonian. The temperature of the heat sources at
time $t$ is $T_{1/2}(t) = T_{1/2}(0)\lambda^{-c t}$, where
$T_{1/2}(0)$ is the initial temperature of the heat sources, and
$\alpha$ is a constant.

Solving the ground state of the Hamiltonian of the Surface Programmable Material from the non-deterministic direction is at least as hard as solving the circuit satisfiability problem, which is NP-complete~\cite{Moore2011}.

@@

\subsubsection{Local cooling test through simulated
annealing}\label{local-cooling-test-through-simulated-annealing}

We will refer to a toy-model limit: $\Delta = 1$ and $U = \infty$.
What's more, we introduced \textbf{``Energy Gradient''} to the gadget to
provide directionality for the simulated annealing. The hamiltonian for
a m-layers automaton now change the form into:

\[
H = \sum_{|\vec r_i - \vec r_j|\leq 2} U \hat n_i \hat n_j  - \Delta \sum_{i,k|\text{vertice i belongs to layer k}} w_i \lambda^{m-k} \hat n_i
\]

The last term of our new hamiltonian represent the \textbf{``Energy
Gradient''}. For vertice $i$ in the k-th layer along the computation
direction, we reset its weight to $w_i\lambda^{m-k}$.

From an intuitive perspective, for layers where the thermal energy
exceeds the energy required to flip the nodes in the current layer,
simulated annealing always filps them randomly; for layers where the
thermal energy is lower than the energy required to flip the nodes in
the current layer, simulated annealing tends to maintain their
configuration; for layers where the thermal energy is just comparable to
the flipping energy, simulated annealing executes the corresponding
cooling process.

Hence, we can simply set the discrete annealing tempretures as
$T(k) = T_0 \eta^k$, where $\eta < 1$ and $\eta ^R = \lambda$.
Here $R$ represent the number of the cooling iterations performed for
a given layer.

Similar to what we did in QAA, firstly, we set the weights of the
inputs/outputs vertices to $\infty$ (NOTE: by removing vertices) for
testing doing computation along deterministic
direction/non-deterministic direction. Then we compare the probability
of successfully finding the corresponding ground state under certain
$T_0, R, \lambda$ and $\eta$. \textbf{We find that doing computation
along the non-deterministic direction is harder than the other one}

Next, we believe that for each layers's cooling process, we are
essentially calculating a probability transition matrix $P(T,k)$,
where $P(T,k)_{outputm, inputm}$ represent the probability that, given
the input vertices state is $inputm$, the cooling process sets the
output vertices state to $outputm$.

Thanks to the translational invariant structure of our gadget and
cooling process, we believe the matrix $P(T,k)$ is independent from
$T$ and $k$, which give us an intuitation that if the failure
probability of each layer's cooling process is $F$, the total success
probability is $(1-F)^m$. \textbf{We test this in a 4-single-gadget
per layer automaton and find this suit well.}

Finally, we evaluate the error probability v.s. run time in a single
layer 4-gadget automaton. Result listed as follows.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/error vs runtime.png}
\caption{Alt text}
\end{figure}

\subsubsection{Estimation of the computing time}\label{estimation-of-the-computing-time}

Let the temperature of the $k$-th layer at time $t$ be
$T(t, k) = T \lambda^{ct - k}$, where $T$ is the initial
temperature, $c$ is a constant, and $\lambda < 1$ is a constant. At
any given time $t$, we denote the subset of atoms at depth
$-\frac{W}{2} < ct - k < \frac{W}{2}$ as the \textbf{active zone},
where $W$ is the width of the sliding window such that
$e^{-\Delta E_{max} /\lambda^{W/2}} = \epsilon \ll 1$. The active zone
is the region where non-trivial computation occurs. The atoms outside
the active zone are either frozen or completely randomized. Clearly,
$W$ asymptotically scales as
$(1-\lambda)^{-1} \log(-\frac{1}{\log\epsilon})$
(NOTE:$W \sim \log_{\lambda}(-\frac{1}{\log{\epsilon}}) = \frac{\log(-\frac{1}{\log{\epsilon}})}{\log(\lambda)} \approx (1 - \lambda)^{-1}\log(-\frac{1}{\log \epsilon})$,
the last term exists when $\lambda \lessapprox 1$).

We consider thermalizing the system in units of $W$. $\epsilon$ is
the error probability of each unit of time, which should scale as
$\epsilon \sim\left(\frac{m}{W}\right)^{-1}$, where $m$ is the total
number of time steps.

@@

Under the assumption that
$\left(1-\frac{\lambda_2(P)}{\lambda_1(P)}\right)\sim e^{-W}$, we have
$t_{\text{th}} \sim e^{(1-\lambda)^{-1}}\log^2(\epsilon^{-1})$. The
total time for the computation is
$t_{\text{total}} \sim \left(\frac{m}{W}\right)\log^2(\frac{m}{W}) e^{(1-\lambda)^{-1}}$.

\subsubsection{Temperature gradient and Energy
gradient}\label{temperature-gradient-and-energy-gradient}

Temperature gradient and energy gradient is indeed equal. One can simply
make the following modification to the transition probability in the
markov process and find the equality.

@@

The time complexity in $\Lambda$ form is apparently
$T_{total} \sim \frac{m}{W}\log^2(\frac{m}{W})e^{(\Lambda-1)^{-1}}$

The reason that we did this transform is that classical simulated
annealing could only generate heat-equilibrilium Boltzmann Distribution,
where the temperature must remain invariant across space.

Time per computation scales as $E \sim \log^2 \frac{m}{W}$.

One should notice that we used a important hypothesis that
$(1-\frac{\lambda_2}{\lambda_1}) \sim e^{-W}$, which means that the
spectral gap of the transition matrix has nothing to do with the width
of a single layer. Should this exists? (\textbf{NOTE: should do some
numerical experiments})

@@

Inspired by the way clothes are ironed, we conceived the idea of slowly
scanning a wave-shaped temperature curve from one end to the other. This
method is widely used in studying the heating-recrystallization
properties of real materials~\cite{Zhang2014}.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/zhang2014.png}
\caption{Alt text}
\end{figure}

There are various specific choices for the wave packet, such as
power-law, exponential, and others. In this context, we primarily focus
on a exponential wave packet, which can be directly linked to the
concept of temperature gradient.

@@

We also conducted a numerical simulation on the wave packet model with
the same condition $\Lambda = 1.3$ and $width = 15$ and fitted the
sweep time against $m\log^2(m)$. Result is listed as follows.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/toy_model_wave_packet_time_vs_depth_1_3.png}
\caption{Alt text}
\end{figure}

@@

The classical simulated annealing can only deal with the situation that
the temperature is space-invariant. Although there are many ways to
simulated a real-world system with temperature gradient, such as
Molecular Dynamics~\cite{Bai2015, Deng2006}, but
these methods mainly depend on dynamically adjusting the atomic
velocities to achieve a temperature gradient, which can't be directly
used in our model.

@@

We conducted numerical experiments under the condition
$\Lambda = 1.3, n = 15$, and fitted each sweep time against
$m \log^2(m)$. Where $m$ represents the number of the single layers,
$n$ represents the width per single layer.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/toy_model_gradient_time_vs_depth_1_3.png}
\caption{Alt text}
\end{figure}

@@

\begin{table}
    \centering
\begin{tabular}{|c|c|c|c|c|}
\hline
input 1 & input 2 & input 3 & output & ancilla \\
\hline
$\downarrow$ & $\downarrow$ & $\downarrow$ & $\downarrow$ & $?$ \\
$\downarrow$ & $\downarrow$ & $\uparrow$ & $\uparrow$ & $?$ \\
$\downarrow$ & $\uparrow$ & $\downarrow$ & $\uparrow$ & $?$ \\
$\downarrow$ & $\uparrow$ & $\uparrow$ & $\uparrow$ & $?$ \\
$\uparrow$ & $\downarrow$ & $\downarrow$ & $\downarrow$ & $?$ \\
$\uparrow$ & $\downarrow$ & $\uparrow$ & $\uparrow$ & $?$ \\
$\uparrow$ & $\uparrow$ & $\downarrow$ & $\uparrow$ & $?$ \\
$\uparrow$ & $\uparrow$ & $\uparrow$ & $\downarrow$ & $?$ \\
\hline
\end{tabular}
\caption{The eight degenerate ground states of the spin-glass model to implement the $110$ rule.}
\end{table}

@@

\newpage
\appendix

>>>>

\subsection{Cooling computation with direction}\label{sec:direction}

@@

The process of heating and annealing the entire material does not have computational capability, as the system lacks logical directionality.

Logical directionality needs to be introduced through an orderly annealing method. 
Similar to the localized heating-annealing approach that catalyzes the growth of ordered crystals, we introduced a temperature gradient annealing scheme.

More specifically, the computation, with computing direction from in-surface to out-surface, contains the following steps: (1). Initialize the in-surface configuration. By removing
some atoms on the in-surface. (2). Connect the in-surface/out-surface to
external heat sources at temperature $T_1 < T_2$, respectively. We
also require that the energy gap between the ground state and the first
excited state of the Hamiltonian to be $\Delta E< T_1$. (3). Lower the
temperature of the heat sources ``slowly'' to cool the system to the
ground state of the Hamiltonian. The temperature of the heat sources at
time $t$ is $T_{1/2}(t) = T_{1/2}(0)\lambda^{-c t}$, where
$T_{1/2}(0)$ is the initial temperature of the heat sources, and
$\alpha$ is a constant.

The computing direction could be reversed by swaping the temperature of the heat sources connected in-surface and out-surface.
However, solving the ground state from the non-deterministic direction is at least as hard as solving the circuit satisfiability problem, which is NP-complete~\cite{Moore2011}.

@@

\subsubsection{Estimation of the computing time along computation direction}\label{estimation-of-the-computing-time}

We fixed the width of each layer and denote $m$ as the number of layers in this material.

Let the temperature of the $k$-th layer at time $t$ be
$T(t, k) = T \lambda^{ct - k}$, where $T$ is the initial
temperature, $c$ is a constant, and $\lambda < 1$ is a constant. At
any given time $t$, we denote the subset of atoms at depth
$ct-\frac{W}{2}\le k \le ct + \frac{W}{2}$ as the \textbf{active zone},
where $W$ is the width of the sliding window such that
$e^{-\Delta E_{max} /\lambda^{W/2}} = \epsilon \ll 1$. The active zone
is the region where non-trivial computation occurs. The atoms outside
the active zone are either frozen or completely randomized.

Clearly, $W$ asymptotically scales as $\frac{\log(-\frac{1}{\log{\epsilon}})}{\log(\lambda)}$. The situations we care about
are those $\lambda \lessapprox 1$, so one can rewrite $W$ as $W \approx (1-\lambda)^{-1} \log(-\frac{1}{\log\epsilon})$.

We consider thermalizing the system in units of $W$. $\epsilon$ is
the error probability of each unit of time, which should scale as
$\epsilon \sim\left(\frac{m}{W}\right)^{-1}$.

@@

Under the assumption that
$\left(1-\frac{\lambda_2(P)}{\lambda_1(P)}\right)\sim e^{-W}$, we have
$t_{\text{th}} \sim e^{(1-\lambda)^{-1}}\log^2(\epsilon^{-1})$. The
total time for the computation should be
$t_{\text{total}} \sim \left(\frac{m}{W}\right)\log^2(\frac{m}{W}) e^{(1-\lambda)^{-1}}$.

\subsubsection{Estimation of the computing time along input memory direction}\label{Estimation-of-the-computing-time-along-input-memory-direction}

We fixed the number of layers and denote $n$ as the width of each layer in this material.

However, we fail to derive an explicitly expression for the time complexity in this direction. 
Numerical result shows that computing complexity is linear to $n$. Thus a total time
for the computation should scale as $nm\log^2(m)e^{(1-\lambda)^{-1}}$

\subsubsection{Testing model through equilibrilium SA}\label{temperature-gradient-and-energy-gradient}

We used simulated annealing to mimic the heating-annealing process.
However, classical simulated annealing can only handle the situation
where the temperature is space-invariant. Thus, a scheme change is needed.

One can simply
make the following modification to the transition probability in the
markov process and return to the situation, called energy-gradient scheme, where the temperature is space-invariant.

@@

Where we denote $\Lambda = \frac{1}{\lambda}$ and the time complexity in $\Lambda$ form is apparently
$T_{total} \sim \frac{m}{W}\log^2(\frac{m}{W})e^{(\Lambda-1)^{-1}}$

@@

Inspired by the way clothes are ironed, we conceived an idea called wave-packet scheme, that slowly
scanning a wave-shaped temperature curve from one end to the other. This
method is widely used in studying the heating-recrystallization
properties of real materials~\cite{Zhang2014}.

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{../notes/images/zhang2014.png}
\caption{Method to scan the wave packet}
\end{figure}

There are various specific choices for the wave packet, such as
power-law, exponential, and others. In this context, we primarily focus
on a exponential wave packet, which can be directly linked to the
concept of energy gradient.

@@

@@

The classical simulated annealing can only deal with the situation that
the temperature is space-invariant. Although there are many ways to
simulated a real-world system with temperature gradient, such as
Molecular Dynamics \cite{Bai2015, Deng2006}, but
these methods mainly depend on dynamically adjusting the atomic
velocities to achieve a temperature gradient, which can't be directly
used in our model.

@@

We conducted numerical experiments both for energy-gradient and wave-packet scheme, under the condition
$\Lambda = 1.3, n = 15$ , and fitted each sweep time against
$m \log^2(m)$. Where $m$ represents the number of the single layers,
$n$ represents the width per single layer.

\begin{figure}[h]
    \centering
    \begin{minipage}{0.48\linewidth} 
        \centering
        \includegraphics[width=\linewidth]{../notes/images/toy_model_gradient_time_vs_depth_1_3.png} 
        \caption{Energy-gradient scheme result}
        \label{fig:image1}
    \end{minipage}
    \begin{minipage}{0.48\linewidth} 
        \centering
        \includegraphics[width=\linewidth]{../notes/images/toy_model_wave_packet_time_vs_depth_1_3.png}
        \caption{Wave-packet scheme result}
        \label{fig:image2}
    \end{minipage}
\end{figure}

Another numerical experiment are conducted to determine the time complexity of the computation along the input memory direction.

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/toy_model_time_vs_width_wave_packet.png}
    \caption{Time v.s. width of the material}
\end{figure}

This indeed informs that the time scale along the memory direction is linear.

@@

\begin{table}[H]
    \centering
\begin{tabular}{|c|c|c|c|c|}
\hline
input 1 & input 2 & input 3 & output & ancilla \\
\hline
$\downarrow$ & $\downarrow$ & $\downarrow$ & $\downarrow$ & $?$ \\
$\downarrow$ & $\downarrow$ & $\uparrow$ & $\uparrow$ & $?$ \\
$\downarrow$ & $\uparrow$ & $\downarrow$ & $\uparrow$ & $?$ \\
$\downarrow$ & $\uparrow$ & $\uparrow$ & $\uparrow$ & $?$ \\
$\uparrow$ & $\downarrow$ & $\downarrow$ & $\downarrow$ & $?$ \\
$\uparrow$ & $\downarrow$ & $\uparrow$ & $\uparrow$ & $?$ \\
$\uparrow$ & $\uparrow$ & $\downarrow$ & $\uparrow$ & $?$ \\
$\uparrow$ & $\uparrow$ & $\uparrow$ & $\downarrow$ & $?$ \\
\hline
\end{tabular}
\caption{The eight degenerate ground states of the spin-glass model to implement the $110$ rule.}
\end{table}

@@

\appendix
